Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-06-13 thirdPaper/draft/typeConeAssociahedra.tex
layout

% formating the table of contents
\makeatletter
\def\l@section{\@tocline{1}{2pt}{0pc}{}{}}
\makeatother
\let\oldtocpart=\tocpart
\renewcommand{\tocpart}[2]{\bf\large\oldtocpart{#1}{#2}}
\let\oldtocsection=\tocsection
\renewcommand{\tocsection}[2]{\bf\oldtocsection{#1}{#2}}

@@

This paper focuses on a surprising construction of the associahedron that recently appeared in mathematical physics.
Motivated by the prediction of the behavior of scattering particles, N.~Arkani-Hamed, Y.~Bai, S.~He, and G.~Yan recently described in~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan} the kinematic associahedron.
It is a class of polytopal realizations of the classical associahedron obtained as sections of a high-dimensional positive orthant with well-chosen affine subspaces.
This construction provides a large degree of freedom in the choice of the parameters defining these affine subspaces, and actually produces all polytopes whose normal fan is affinely equivalent to that of J.-L.~Loday's associahedron~\cite{Loday} (see \cref{subsec:typeConeAsso}).
These realizations were then extended by V.~Bazier-Matte, G.~Douville, K.~Mousavand, H.~Thomas and E.~Y\i ld\i r\i m~\cite{BazierMatteDouvilleMousavandThomasYildirim} in the context of finite type cluster algebras using tools from representation theory of quivers.
More precisely, they fix a finite type cluster algebra~$\cal{A}$ and consider the real euclidean space~$\R^\cal{V}$ indexed by the set~$\cal{V}$ of cluster variables of~$\cal{A}$.
Starting from an acyclic initial seed~$\seed_\circ$, they consider a set~$\cal{M}$ of mesh mutations and compute the corresponding linear dependences between the $\b{g}$-vectors, which can be interpreted as linear spaces in~$\R^\cal{V}$.
Finally, they perturb these mesh linear spaces by a collection~$\b{\ell} \in \R_{>0}^{\cal{M}}$ of positive parameters and intersect the resulting perturbed mesh affine spaces with the positive orthant~$\R_{\ge 0}^\cal{V}$ (see \cref{subsec:typeConeCA}).
They show that the resulting polytope is always a generalized associahedron, and that its normal fan is the $\b{g}$-vector fan of~$\cal{A}$ with respect to the initial acyclic seed~$\seed_\circ$.
An implicit by-product of their construction is that the space of all polytopal realizations of the $\b{g}$-vector fan of~$\cal{A}$ with respect to the initial acyclic seed~$\seed_\circ$ is a simplicial cone.

@@

\begin{remark}
\label{rem:exchangeablePairsNKC}
In view of \cref{prop:exchangeablePairsNKC}, it is tempting to look for a characterization of the exchangeable pairs~$\omega, \omega'$ using the kisses between~$\omega$ and~$\omega'$.
This question was discussed in~\cite[Sect.~9]{BrustleDouvilleMousavandThomasYildirim}.
However, as illustrated for instance in the non-kissing complex of \cref{fig:nonkissingFans}\,(right),
\begin{itemize}
\item two exchangeable walks may kiss along more than one string (only one is distinguished),
\item two non-exchangeable walks can kiss along more than one distinguishable string,
\item two walks that kiss along a single distinguishable string are not always exchangeable,
\item not all strings are distinguishable.
\end{itemize}
In \cref{subsubsec:simplicialTypeConeNKC}, we will restrict to a situation that avoids all these patologies.
%\vincent{Do we still have these problems when the gentle algebra is brick and/or without oriented $2$-cycle?}
\end{remark}

@@

\begin{figure}[t]
	\capstart
	\begin{adjustbox}{center}
		\begin{tabular}{c@{\;}c@{\;}c@{\;}c}
			permutahedron &
			associahedron &
			cyclohedron &
			stellohedron \\
			\includegraphics[scale=.6]{permutahedron} &
			\includegraphics[scale=.6]{associahedron} &
			\includegraphics[scale=.6]{cyclohedron} &
			\includegraphics[scale=.6]{stellohedron} \\
			\includegraphics[scale=.6]{permutahedronTubings} &
			\includegraphics[scale=.6]{associahedronTubings} &
			\includegraphics[scale=.6]{cyclohedronTubings} &
			\includegraphics[scale=.6]{stellohedronTubings} \\[.1cm]
			complete graph &
			path &
			cycle &
			star
		\end{tabular}
	\end{adjustbox}
	\caption{Some classical families of polytopes as graph associahedra. Illustration from~\cite{MannevillePilaud-compatibilityFans}.}
	\label{fig:specialGraphAssociahedra}
\end{figure}

@@

\subsubsection{Nested complex and nested fan of a graph}

@@

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[width=\textwidth]{exmNested}}
	\caption{Some incompatible tubes (left and middle), and a maximal tubing (right).}
	\label{fig:exmNested}
\end{figure}

@@

\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=.55]{nestedFans}}
	\caption{Two nested fans. As the fans are $3$-dimensional, we intersect them with the sphere and stereographically project them from the direction~$(-1,-1,-1)$.}
	\label{fig:nestedFans}
\end{figure}

@@

\begin{corollary}
\label{corollary:UERPforCAproof}
\emph{(\cref{prop:uniqueExchangePropertyCA})}
Let~$\B_\circ$ be any finite type exchange matrix, and let~$\mathcal{A}(\B_\circ)$ be the associated cluster algebra without coefficients.
Then, for any exchangeable cluster variables~$x$ and~$x'$, the linear dependence $\sum_{y \in \cluster \cup \cluster'} \coefficient[{\b{g}_y}][\cluster][\cluster'] \, \b{g}_y = 0$ only depends on the pair~$(x,x')$ and not on the specific choice of clusters~$\cluster$ and~$\cluster'$ containing~$x$ and~$x'$ respectively and such that~${\cluster\setminus\{x\}=\cluster'\setminus\{x'\}}$.
\end{corollary}

@@

The Auslander--Reiten quiver of the module category~$\MOD(\Lambda)$ can be obtained by applying the functor~$\Hom{}(\Lambda, -)$.

@@

One can check directly that the condition of \cref{theo::brick-algebra-condition} is not satisfied.
For example, the module~$\bsm2\\1\esm$ is~$\tau$-rigid since~$\tau(\bsm2\\1\esm) = \bsm3\\1\esm$, while~$\Hom{\Lambda}\left(\bsm2\\1\esm,\tau^2(\bsm2\\1\esm)\right) = \Hom{\Lambda}(\bsm2\\1\esm,\bsm1\\2\esm) \ne 0$.
\end{example}

>>>>

% formating the table of contents
\setcounter{tocdepth}{4}
\makeatletter
\def\l@section{\@tocline{1}{2pt}{0pc}{}{}}
\makeatother
\let\oldtocpart=\tocpart
\renewcommand{\tocpart}[2]{\bf\large\oldtocpart{#1}{#2}}
\let\oldtocsection=\tocsection
\renewcommand{\tocsection}[2]{\bf\oldtocsection{#1}{#2}}
\let\oldtocsubsubsection=\tocsubsubsection
\renewcommand{\tocsubsubsection}[2]{\quad\oldtocsubsubsection{#1}{#2}}

@@

This paper focuses on a surprising construction of the associahedron that recently appeared in mathematical physics.
Motivated by the prediction of the behavior of scattering particles, N.~Arkani-Hamed, Y.~Bai, S.~He, and G.~Yan recently described in~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan} the kinematic associahedron.
It is a class of polytopal realizations of the classical associahedron obtained as sections of a high-dimensional positive orthant with well-chosen affine subspaces.
This construction provides a large degree of freedom in the choice of the parameters defining these affine subspaces, and actually produces all polytopes whose normal fan is affinely equivalent to that of J.-L.~Loday's associahedron~\cite{Loday} (see \cref{subsec:typeConeAsso}).
These realizations were then extended by V.~Bazier-Matte, G.~Douville, K.~Mousavand, H.~Thomas and E.~Y\i ld\i r\i m~\cite{BazierMatteDouvilleMousavandThomasYildirim} in the context of finite type cluster algebras using tools from representation theory of quivers.
More precisely, they fix a finite type cluster algebra~$\cal{A}$ and consider the real euclidean space~$\R^\cal{V}$ indexed by the set~$\cal{V}$ of cluster variables of~$\cal{A}$.
Starting from an acyclic initial seed~$\seed_\circ$, they consider a set~$\cal{M}$ of mesh mutations and compute the corresponding linear dependences between the $\b{g}$-vectors, which can be interpreted as linear spaces in~$\R^\cal{V}$.
Finally, they perturb these mesh linear spaces by a collection~$\b{\ell} \in \R_{>0}^{\cal{M}}$ of positive parameters and intersect the resulting perturbed mesh affine spaces with the positive orthant~$\R_{\ge 0}^\cal{V}$ (see \cref{subsec:typeConeCA}).
They show that the resulting polytope is always a generalized associahedron, and that its normal fan is the $\b{g}$-vector fan of~$\cal{A}$ with respect to the initial acyclic seed~$\seed_\circ$.
An implicit by-product of their construction is that the space of all polytopal realizations of the $\b{g}$-vector fan of~$\cal{A}$ with respect to the initial acyclic \mbox{seed~$\seed_\circ$ is a simplicial cone}.

@@

\begin{remark}
\label{rem:exchangeablePairsNKC}
\enlargethispage{.3cm}
In view of \cref{prop:exchangeablePairsNKC}, it is tempting to look for a characterization of the exchangeable pairs~$\omega, \omega'$ using the kisses between~$\omega$ and~$\omega'$.
This question was discussed in~\cite[Sect.~9]{BrustleDouvilleMousavandThomasYildirim}.
However, as illustrated for instance in the non-kissing complex of \cref{fig:nonkissingFans}\,(right),
\begin{itemize}
\item two exchangeable walks may kiss along more than one string (only one is distinguished),
\item two non-exchangeable walks can kiss along more than one distinguishable string,
\item two walks that kiss along a single distinguishable string are not always exchangeable,
\item not all strings are distinguishable.
\end{itemize}
In \cref{subsubsec:simplicialTypeConeNKC}, we will restrict to a situation that avoids all these patologies.
%\vincent{Do we still have these problems when the gentle algebra is brick and/or without oriented $2$-cycle?}
\end{remark}

@@

\hvFloat[floatPos=p, capWidth=h, capPos=r, capAngle=90, objectAngle=90, capVPos=c, objectPos=c]{figure}
{
		\begin{tabular}{c@{\;}c@{\;}c@{\;}c}
			permutahedron &
			associahedron &
			cyclohedron &
			stellohedron \\[.2cm]
			\includegraphics[scale=.93]{permutahedron} &
			\includegraphics[scale=.93]{associahedron} &
			\includegraphics[scale=.93]{cyclohedron} &
			\includegraphics[scale=.93]{stellohedron} \\[.1cm]
			\includegraphics[scale=.93]{permutahedronTubings} &
			\includegraphics[scale=.93]{associahedronTubings} &
			\includegraphics[scale=.93]{cyclohedronTubings} &
			\includegraphics[scale=.93]{stellohedronTubings} \\[.2cm]
			complete graph &
			path &
			cycle &
			star
		\end{tabular}
}
{Some classical families of polytopes as graph associahedra. Illustration from~\cite{MannevillePilaud-compatibilityFans}.}
{fig:specialGraphAssociahedra}

%\begin{figure}[t]
%	\capstart
%	\begin{adjustbox}{center}
%		\begin{tabular}{c@{\;}c@{\;}c@{\;}c}
%			permutahedron &
%			associahedron &
%			cyclohedron &
%			stellohedron \\
%			\includegraphics[scale=.6]{permutahedron} &
%			\includegraphics[scale=.6]{associahedron} &
%			\includegraphics[scale=.6]{cyclohedron} &
%			\includegraphics[scale=.6]{stellohedron} \\
%			\includegraphics[scale=.6]{permutahedronTubings} &
%			\includegraphics[scale=.6]{associahedronTubings} &
%			\includegraphics[scale=.6]{cyclohedronTubings} &
%			\includegraphics[scale=.6]{stellohedronTubings} \\[.1cm]
%			complete graph &
%			path &
%			cycle &
%			star
%		\end{tabular}
%	\end{adjustbox}
%	\caption{Some classical families of polytopes as graph associahedra. Illustration from~\cite{MannevillePilaud-compatibilityFans}.}
%	\label{fig:specialGraphAssociahedra}
%\end{figure}

@@

\clearpage
\subsubsection{Nested complex and nested fan of a graph}

@@

\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[width=\textwidth]{exmNested}}
	\caption{Some incompatible tubes (left and middle), and a maximal tubing (right).}
	\label{fig:exmNested}
\end{figure}

@@

\begin{figure}[b]
	\capstart
	\centerline{\includegraphics[scale=.55]{nestedFans}}
	\caption{Two nested fans. As the fans are $3$-dimensional, we intersect them with the sphere and stereographically project them from the direction~$(-1,-1,-1)$.}
	\label{fig:nestedFans}
\end{figure}

@@

\begin{corollary}
\label{corollary:UERPforCAproof}
\emph{(\cref{prop:uniqueExchangePropertyCA})}
Let~$\B_\circ$ be any finite type exchange matrix, and let~$\mathcal{A}(\B_\circ)$ be the associated cluster algebra without coefficients.
Then, for any exchangeable cluster variables~$x$ and~$x'$, the linear dependence $\sum_{y \in \cluster \cup \cluster'} \coefficient[{\b{g}_y}][\cluster][\cluster'] \, \b{g}_y = 0$ only depends on the pair~$(x,x')$ and not on the specific choice of clusters~$\cluster$ and~$\cluster'$ containing~$x$ and~$x'$ respectively and such that~${\cluster \ssm \{x\} = \cluster' \ssm \{x'\}}$.
\end{corollary}

@@

\medskip
The Auslander--Reiten quiver of the module category~$\MOD(\Lambda)$ can be obtained by applying the functor~$\Hom{}(\Lambda, -)$.

@@

\medskip
One can check directly that the conditions of \cref{theo::brick-algebra-condition} is not satisfied.
For example, the module~$\bsm2\\1\esm$ is~$\tau$-rigid since~$\tau(\bsm2\\1\esm) = \bsm3\\1\esm$, while~$\Hom{\Lambda}\left(\bsm2\\1\esm,\tau^2(\bsm2\\1\esm)\right) = \Hom{\Lambda}(\bsm2\\1\esm,\bsm1\\2\esm) \ne 0$.
\end{example}
